---
title: "DISNEP simulation"
author: "ZHUOHUI Liang"
date: "7/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(MASS)
library(SMUT)
library(foreach)
library(parallel)
library(doParallel)
library(tidyverse)

set.seed(123123)

knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)
```

# Simulation Settings

```{r give_setting}
settings = tibble(
  id = list(1:50),
  scenario = 1:5,
  n_signal = rep(50,5),
  n_noise  = rep(1950,5),
  strong_signal_mu = c(10,10,10,11,12),
  median_signal_mu = c(9,9,9,10,11),
  noise_mu = rep(8,5),
  strong_corr = rep(0.1,5),
  median_corr = rep(0.02,5),
  noise_corr = rep(0.01,5),
  sigma = c(3,2,1,1,1),
  prob_jump = rep(0.5,5)
) %>% 
  filter(scenario %in% 1:5)


knitr::kable(t(settings %>% select(-id)),
             caption = "Simulation settings")

settings = settings %>% 
  filter(scenario %in% 1:5) %>% 
  select(-scenario) %>% 
  unnest(id) %>% 
  mutate(id = order(id))
```

   For each simulation, there will be two data/matrix simulated, namely gene-gene-interaction network and gene-expression data. Each simulation we will compare t-test, GeneWanderer( random walk with restart on gene-gene-interaction network) and Disnep( random walk with restart on disease-enhanced-gene-gene-interaction network[Ruan,2019])
   
## Gene-gene-interaction network

   In each simulation, network is sample from existing database, we have homo sapiens(human) gene-gene(protein-protein) interaction network(PPI) downloaded from V.11 STRING. Since Genewanderer only consider gene-gene-interaction score > 0.4, we use Cytoscape to pull STRING node information from STRING containing only score > 0.4 and map Gene name to PPI.
   
   And we set the gene from DISGENET V7 renal carcinoma disease(C1378703) as disease/signal gene  list and the rest of the gene not in DISGENET list as noise gene.
   
   We sample `r settings$n_signal[[1]]` gene from disease gene list and `r settings$n_noise[[1]]` as noise gene. We keep the edge of the gene from STRING PPI network and set all missing value as 0. And we order the first `r settings$n_signal[[1]]` gene as signal gene just for convenient.
   

## Gene Expression network

  In each simulation, we generate 300 case and 50 control sample. 
  
### Case

  The case set to have `r settings$n_signal[[1]]/2` strong signal gene, which each gene's mean is generated from N($\mu_1$,1), and use this mean to generate gene expression level. case has another `r settings$n_signal[[1]]/2` median signal gene, which each gene's mean is generated from N($\mu_2$,1). the rest `r settings$n_noise[[1]]` gene are set as noise gene, which's mean is generate from N(8,1).
  
  The correlation between strong signal is sampled from U(0,0.1) and correlation between median signal is sampled from U(0,0.02) and correlation between noise is sampled from U(0,0.01). 
  
\begin{equation}
Cov = sigma^2 *
\begin{pmatrix}
1 & \rho_{12} & 0 &0\\
\rho{21} & 1 & \rho{23} & 0 \\
0 & \rho{32} & 1 & \rho{34} \\
0 & 0 & \rho{43} & 1\\
\end{pmatrix}
\end{equation}

where $\rho_{12} = \rho_{21}^t$ is the triagle block sample from U(0,0.1) and vise versa for $\rho_{23},\rho{34}$

## control

 All genes in control is set to be noise and its mean generate from N(8,1) and correlation sample from U(0,0.01)
 


all gene expression is generated from multivariate normal, with generated means and covariate as mention above.

  


```{r simulation,eval=T}

source("R/simulation.R")

system("rm cache/**/*.Rdata -r")

#mclapply(1:nrow(settings),
#         FUN = function(x) {pmap(settings[x,],simulation)},
#         mc.cores = 4
#         )

cl = parallel::makePSOCKcluster(3)

doParallel::registerDoParallel(cl)

result = foreach(
  i = 1:nrow(settings),
  .errorhandling = "pass",
  .packages = c(
    "SNFtool",
    "MASS",
    "SMUT",
    "Matrix",
    "readr",
    "dplyr",
    "tidyverse"
  )
) %dopar% pmap(settings[i, ], simulation)

stopCluster(cl)

```




```{r data_example}
result = readRDS(list.files("cache/simulation/",full.names = T,pattern = "Rdata")[[1]])

a = as.matrix(result$gene_int[[1]])

heatmap(
  a[1:500, 1:500],
  Rowv = NA,
  Colv = NA,
  col = topo.colors(8),
  main = "Interaction first 500(included signals)"
)

a = colMeans(result$gene_expr[[1]])


plot(1:500, a[1:500], main = paste("gene expression with mu", paste0(result[, 3:5], collapse = "-")))
points(1:result$n_signal[[1]], a[1:result$n_signal[[1]]], col = "red", add = T)


a = abs(cor(result$gene_expr[[1]]))

diag(a) = 0

heatmap(
  t(a)[1:500, 1:500],
  Rowv = NA,
  Colv = NA,
  col = topo.colors(8),
  main = "Correlation first 500(included signals)"
)

invisible(gc())
```

```{r auc}
result = foreach(file = list.files("cache/result/",full.names = T,pattern = "Rdata"),
                  .combine = bind_rows,
                 .errorhandling = "remove") %do% readRDS(file)

result = result %>%
  unite("mu_sigma", c(strong_signal_mu:noise_mu, sigma)) %>%
  select(-ends_with("corr")) %>% 
  mutate(
    disease_gene = map2(n_signal,n_noise,  ~ c(rep(1,.x),rep(0,.y))),
    score = map(score,~arrange(.x,gene)$score),
    auc = map2(score,disease_gene,~mltools::auc_roc(.x/max(.x),.y,returnDT = T))
    ) %>% 
  unnest(auc) %>% 
  mutate(top_k = map2_dbl(score,Pred,~sum((.x/max(.x))>=.y)))

result %>%
  group_by(method, mu_sigma,top_k) %>%
  summarise(sens = mean(CumulativeTPR),
            spec = mean(1-CumulativeFPR),
            fpr = mean(CumulativeFPR)) %>%
  ggplot(aes(x = fpr, y = sens, color = method)) +
  geom_step()+
  geom_abline(slope = 1,alpha = 0.25)+
  facet_wrap(.~mu_sigma)+
  labs(title = "ROC")

result %>%
  filter(CumulativeFPR<=0.1) %>% 
  group_by(id,method, mu_sigma) %>%
  summarise(auc = max(CumulativeArea)) %>% 
  ggplot(aes(x = method, y = auc, color = method)) +
  geom_boxplot()+
  facet_wrap(.~mu_sigma,scales = "free_y")+
  labs(title = "AUC(FPR<0.1)")


result %>%
  filter(CumulativeFPR<=0.1) %>% 
  group_by(id,method, mu_sigma) %>%
  summarise(auc = max(CumulativeArea)) %>% 
  group_by(method,mu_sigma) %>% 
  summarise(mean_auc = mean(auc)) %>% 
  pivot_wider(names_from = method,
              values_from = mean_auc) %>% 
  knitr::kable(caption = "mean AUC(FPF<0.1)",
               digits = 3)
```




